---
title: "GSS challange"
author: "Jessica Villa"
date: "Summer 2022"
output: html_document
---







```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
GSS <-read.dta(file="GSS2018.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,race,cappun)
mydata <- na.omit(mydata)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Is attitude toward capitol punishment (cappun) independent of race (race)?
If not, then whose view differs from whom??

$H_0$: Attitude toward capital punishment is not related to race  

$H_A$: Attitude toward capital punishment is dependent on race  


# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

Both race and cappun are categorical variables so the analysis technique is CAT~CAT; plot is a barplot, summary is cross-table, inference is chi-squared test.  


# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->



## Now with The Most data (fewer variables gives more data points)

```{r}
dd2 <- mydata %>% group_by(race,cappun) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=race,y=count,fill=cappun))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=race,y=prcnt*100,fill=cappun)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

According to the chart the race "white" has an overall greater percentage (73%) of being in favor of cappun, with race "other" holding (55%), and race "black" last with 45% in favor of cappun.  


### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->


According to the number chart, out of 100 race "white" 67.98 are in favor of cappun and 32.02 are opposed to cappun. Out of 100 race "black" 46.29 are in favor of cappun and 53.71 are opposed to cappun. Out of the 100 race "other 55.87 are in favor of cappun and 44.13 are opposed to cappun. In addition, out of the 100 races in favor of cappun; 78.34 are of "white" race, 11.7 are of "black" race, and 9.96 are other. Out of the 100 of races opposed to cappun: 63.24 are white, 23.27 are black, and 13.49 are other.  



## Now with More data

```{r}
table2 <- xtabs(~race + cappun, data=mydata)
rowPerc(table2)
colPerc(table2)
```



## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

 


## More data version

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
#fisherXY(mydata$race,mydata$cappun)
```

The p-value for the chi-squared test was .00000000000002033. The p-value is very small so we reject the null hypotheses. We reject it because both p-values are way smaller than .05. The p-value for black vs other race is 0.02479 so regardless, both p-values given are different. 

```{r}
fisherXY(mydata$cappun,mydata$race)
```
 

# Discussion and Conclusion

Our question was whether or not different races had different opinions on capital punishment. We stated that both values are categorical so we worked around it. We used a bar plot, a table, and a chi-squared test. From the following, we learned that there is a difference. Both our p-values were below .05. The white race was way more in favor of capital punishment compared  to any other race by 14%. Thus, we rejected the null hypothesis due to there being way too much evidence against the claim that there was no correlation between race & their opinions. 

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->


<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->



  
  

  